\plaintitle

### 2017-01-06 drafts/labor/complexity.tex
lit review except for airy's computers done

\clearpage
\subsubsection{\pieceworkpers}

@@

{\topic{Only tasks that could be measured and priced could be completed via piecework
--- a fact which fundamentally limited the potential complexity of piecework.}
When \citeauthor{Brown01041990} investigated
what limited the adoption of piecework in industries that otherwise gravitated toward it
(e.g., railway engineers),
the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
\citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
which used ``efficiency experts'' to develop a ``standard time''
to determine pay for each task at the company informed by
``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
~\cite{10.2307/23702539}.
One might conclude from \citeauthor{10.2307/23702539}'s observations that
complex, creative work
--- which is inherently heterogeneous and difficult to routinize ---
would be unsuitable for piecework.

@@

\clearpage
\subsubsection{\whatchanged}
\begin{comment}
  mangeerial overhead limits, so what's different
  more people can now do complex work without training (more complex)
  parts of management can be automated (more firms)
  cheaper to create the infrastructure (more complex)
\end{comment}

@@

\clearpage
\subsubsection{\implication}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.

>>>>

% \clearpage
\subsubsection{\pieceworkpers}

@@

{\topic{Only tasks that could be measured and priced could be completed via piecework
--- a fact which fundamentally limited the potential complexity of piecework.}
% When \citeauthor{Brown01041990} investigated
% what limited the adoption of piecework in industries that otherwise gravitated toward it
% (e.g., railway engineers),
% the homogeneity of tasks arose as a major contributing factor~\cite{Brown01041990}.
% \citeauthor{10.2307/23702539} concurs via a case study of the Santa Fe Railway,
% which used ``efficiency experts'' to develop a ``standard time''
% to determine pay for each task at the company informed by
% ``thousands of individual operations''% removed ``at the Topeka shop'' because even I had to reach back into the back of my mind to be like ``Oh, yeah, Kansas''.
% ~\cite{10.2307/23702539}.
Earlier we discussed \citeauthor{10.2307/23702539}'s and later \citeauthor{Brown01041990}'s analysis of railway workers;
one might conclude from \citeauthor{10.2307/23702539}'s observations in particular that
complex, creative work
--- which is inherently heterogeneous and difficult to routinize ---
would be unsuitable for piecework~\cite{10.2307/23702539}.
\citeauthor{Brown01041990}'s description of efficiency experts would corroborate this;
``efficiency experts'' can effectively gauge how long known tasks should take, but
would find themselves overwhelmed if they attempted to assess creative work like design,
which can take an arbitrary number of iterations before proceeding to a subsequent step
\ali{should i cite some design research here?}.

@@

% \clearpage
\subsubsection{\whatchanged}
\begin{comment}
  mangeerial overhead limits, so what's different
  more people can now do complex work without training (more complex)
  parts of management can be automated (more firms)
  cheaper to create the infrastructure (more complex)
\end{comment}

@@

% \clearpage
\subsubsection{\implication}
Technology's ability to support human cognition will enable stronger assumptions about workers' abilities, increasing the complexity of crowd work outcomes.
Just as the shift to expert crowdsourcing increased complexity, so too will workers with better tools increase the set of tasks possible.
Beyond this, further improvements would most likely come from replicating the success of narrowly--slicing education for expert work as \citeauthor{hart2013rise} and \citeauthor{grier2013computers} described in their piecework examples
of human computation~\cite{grier2013computers} and drastically reformulating macro--tasks given the constraints of piecework~\cite{hart2013rise}.
To some extent, an argument can be made that
MOOCs and other online education resources
provide crowd workers with the resources that they need, but 
it remains to be seen whether that work will be appropriately valued, let alone
properly interpreted by task solicitors~\cite{aguaded2013mooc}.
If we can overcome this obstacle,
we might be able to empower more crowd workers to do complex work such as engineering and metalworking,
rather than doom them to ``uneducated'' match girl reputations~\cite{10.2307/3827491}.
However, many such experts are already available on platforms such as Upwork, so training may not directly increase the complexity accessible to crowd work unless it makes common expertise more broadly available.

### 2017-01-06 drafts/labor/piecework_lit.tex
lit review except for airy's computers done

\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over the years,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions,
tracing an outline of % a working understanding of
\textit{what piecework is}
alongside piecework's contemporary developments in practice.

@@

% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
particularly driven toward encapsulating the manual labor of farmwork.}
\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
Indeed, employers adopted piece--rate compensation in some aspects and
time--rate compensation in others.
The Rowan premium system,
which essentially paid workers
a base rate for time plus, potentially,
an additional pay dependent on output~\cite{rowan1901premium};
this was just one of several alternatives to categorical time-- and piece--rate renumeration paradigms.
% which will complicate our categorization of cases of piecework.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} additional compensation tied to performance,
workers under this regime were
in some senses ``task--labourers'', but
in other senses ``day--labourers''.

\topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework, % do you like what i did there?
pointing out that
``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~\dots~he knows that all depends on his own diligence and perseverance~\dots~[and] so long as he performs his work to the satisfaction of his master, he is not under that control to which the day--labourer is always subject''.
The argument that ``task--labourers'' enjoy freedom from control crops up in \citeauthor{hughRaynbirdTaskWork}'s and later \citeauthor{rowan1901premium}'s works~\cite{hughRaynbirdTaskWork,rowan1901premium}.

@@

% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.

@@

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
railroad and other industrial workers;
Airy's employment of \textit{human computers};
and domestic and farm work (in particular, the ``match--girls'').}
We'll introduce the facts of these three cases,
beginning with the most familiar case of piecework
--- industrial work, such as assembly line manufacturing ---
working backwards through the \nth{20} and \nth{19} century.
In doing so, we'll trace the history of piecework
while also framing the later analysis of the major research threads we named earlier
(questions toward complexity, decomposition, and relationships, respectively).

\ali{I'm coming out of this with a sort of structure that \textit{loosely} describes piecework chronologically.
Strictly speaking, I'm talking about each of these topics
(let's call them
\textbf{human computation},
\textbf{domestic and farm work}, and
\textbf{industrial work}),
and some of these things overlap with each other to some extent, but
for the most part
Airy and the human computers came in the mid--\nth{19} century,
farm work and the ``domestic'' work (like making matches) came around at the turn of the \nth{20} century, and
the industrial work (railroad workers, the WWII war effort, and the rise of the labor union movement) unfolded from the early \nth{20} century onward.}

\subsubsection{Railroad and other Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}

\msb{I'm feeling very disoriented in this paragraph. The subsection title is about railroads, but we're all fast-forwarded to auto manufacturing?}
\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% This approach would turn out to yield two benefits:
% \item these workers could be productive with significantly less ``on--boarding'' cost
% \item workers' performance could be tracked \& evaluated, and better methods could be disseminated to workers.

\msb{confusing timing; now we're going backwards?}
{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's eponymously named \textit{Fordism}),
piecework in the early and mid--\nth{20} century surged, especially in industrial work.}
Scientific management promised that the careful measurement of workers would yield
higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
While \citeauthor{Brown01041990} points out that
piecework dramatically advanced the instrumented measurement of workers,
in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
using completion times to inform per--task compensation.
% Under this regime, piecework became commonplace in railroad maintenance and a number 
% This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
% railroad maintenance, automobile manufacturing, and later weapons manufacturing.
% \ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

% \topic{With the formalization of scientific management,
% piecework became an important component of the mass manufacturing
% in American factories.}
% \topic{}

\msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
\msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

\msb{This subsection feels like it never delivered on what it promised: a case study of one group of pieceworkers. I thought the goal was to introduce the railroad workers --- who they were, what they did, what their conditions were like.}

% \onlyinsubfile{\clearpage}
\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
\msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
\msb{how is this sentence relevant to the case study?:}
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

\topic{Workers' relationships with employers quickly soured.}
\msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
\msb{Why is Taylor relevant to this case study?}
\ali{I'm just clumsily interleaving the contemporary literature with the events.
I'd like to say ``This is the geist of work in their time'';
is this a bad idea?}
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23
\ali{this has a ``gosh isn't that interesting?'' feeling to it.
should i just leave it out until later when there's something to say about it
(namely, that the mode of work was initially designed to stymie professional growth)?}%
.

>>>>

\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over the years,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions,
tracing an outline of
\textit{what piecework is}
alongside piecework's contemporary developments in practice.

@@

\topic{By \citeyear{hughRaynbirdTaskWork} we find
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay
(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
particularly driven toward encapsulating the manual labor of farmwork.}
\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
Indeed, employers adopted piece--rate compensation in some aspects and
time--rate compensation in others.
The Rowan premium system,
which essentially paid workers
a base rate for time plus, potentially,
an additional pay dependent on output~\cite{rowan1901premium};
this was just one of several alternatives to categorical time-- and piece--rate renumeration paradigms.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} additional compensation tied to performance,
workers under this regime were
in some senses ``task--labourers'', but
in other senses ``day--labourers''.

\topic{It may be worth thinking about piecework through the lens of its \textit{emergent} properties to help understand it.}
\citeauthor{hughRaynbirdTaskWork} argues for the merits of piecework,
pointing out that
``piece work holds out to the labourer an increase of wages as a reward for his skill and exertion~\dots~he knows that all depends on his own diligence and perseverance~\dots~[and] so long as he performs his work to the satisfaction of his master, he is not under that control to which the day--labourer is always subject''.
The argument that ``task--labourers'' enjoy freedom from control crops up in \citeauthor{hughRaynbirdTaskWork}'s and later \citeauthor{rowan1901premium}'s works~\cite{hughRaynbirdTaskWork,rowan1901premium}.

@@

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.

@@

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
railroad and other industrial workers;
Airy's employment of \textit{human computers};
and domestic and farm work (specifically, the ``match--girls'').}
In introducing these cases at a high level,
we'll trace the history of piecework
while also framing the later analysis of the major research threads we named earlier
(questions toward complexity, decomposition, and relationships, respectively).

\begin{comment}
What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)

- Graves: railway companies used ``efficiency experts'' to study how long tasks should take
- Hart: evaluation limits complexity (we can affect that with peer evaluation!)
- Graves: sparks of Scientific Management in Piecework
- organization types are important determinants of piecework viability: lots of types of tasks? bad
  - Hart (I think?): variability in *worker* quality is fine
- Foreman is important
- Worker advocacy groups arose to speak out against piecework

\end{comment}

\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
In the railroad industry and especially on the assembly line
we see many of the mechanisms of work and worker management that later made their way into crowd work.
Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
We'll expose these facets of piecework very superficially in the context of industrial work
to better speak to these research questions in crowd work later.

\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
what followed was a flourishing in the management practices with pieceworkers,
as railway companies struggled to find effective ways
to motivate and evaluate this skilled workforce of engineers.}
\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
finding that they employed ``efficiency experts'' to develop a ``standard time''
to determine pay for each task at the company informed by
``thousands of individual operations''; % ~\cite{10.2307/23702539}.
\citeauthor{10.2307/23702539} goes on to list
some of the roles required to facilitate piecework
in the early \nth{20} century
--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.

\topic{This oversight, while controversial
(especially among workers~\cite{american1921problem}),
paved the way for piecework to grow substantially.}
With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's eponymously named \textit{Fordism}),
piecework in the early and mid--\nth{20} century surged, especially in industrial work.
Scientific management promised that the careful measurement of workers would yield
higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
While \citeauthor{Brown01041990} points out that
piecework dramatically advanced the instrumented measurement of workers,
in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
using completion times to inform per--task compensation,
similarly to efficiency experts in the Santa Fe Railway, but
substantially more precise.

\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% \msb{confusing timing; now we're going backwards?}

% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

\ali{some things I didn't cover:
\begin{itemize}
  \item research finding that task variability is bad for piecework, but worker variability is good
  \item talk about the foreman
  \item worker advocacy groups
\end{itemize}
should I be making a serious effort to get that stuff in here? this section is kind of long already
(and it'll definitely be much longer than the other case studies).}

% \onlyinsubfile{\clearpage}
\subsubsection{Domestic and Farmhand Labor}

\begin{comment}
What did I pull from the threads that are related to domestic/farmwork?

- Graves: sparks of Scientific Management in Piecework, especially starting here
- 19th century: piecework was mostly cottage industry with untrained or informally trained workers
  (unlike industrial metal workers during WWII)
- Brown: Task variability matters
- Clark: pieceworkers work harder, more diligently, etc...
- Riis saw terrible conditions, documented and communicated it to the world

\end{comment}

\topic{The application of piecework to farm work in the late \nth{19} century and
later to manufacturing of small goods, such as garments and matches, at the turn of the \nth{20} century
proved to be a formative period for piecework as we would come to know it.}
Here, then, we'll discuss the implementation of piecework regimes in farms,
in homes where workers assembled clothing and were paid by the \textit{piece} of output, and
finally in the early sweatshops where women made matchsticks under dangerous and even hazardous conditions.

\topic{Farm and field work emerged early on as a rich field for the use of piecework,
perhaps in part due to the relatively low skillset necessary to complete work.}
We discussed earlier that farm work applied the idea of piecework to the farm by
paying workers for things like picking bushels of fruit or bringing to birth animals~\cite{10.2307/2338394}.
\citeauthor{hughRaynbirdTaskWork} argues from another perspective, but along a similar thread, that
the only factor important in piecework was
whether the worker's final product was to the ``satisfaction of his master''~\cite{hughRaynbirdTaskWork}.
Piecework's advantage,
\citeauthor{hughRaynbirdTaskWork} argues,
was this hands--off management of workers.

\topic{Piecework was beginning to find its way into homes near the end of the \nth{19} century.}
Textile manufacturers found that they could deliver fabric to people at their homes,
paying them to sew together clothing.
The manufacturers would later return to retrieve the finished garments,
paying these workers for each \textit{piece} of clothing made.
They could, in principle, assemble as much or as little clothing as they wanted;
the reality was more grim, as
\citeauthor{riisOtherSideLives} documented in \citeyear{riisOtherSideLives}~\cite{riisOtherSideLives}.
He found that
workers endured bleak living conditions and
worked long hours attempting to scrape together a living.

\topic{Piecework played a role in early mass manufacturing during the Industrial Revolution,
as well as the early years of worker collective organization and action.}
At the end of the \nth{19} century,
manufacturers employed women to assemble matchsticks in factories.
These match--makers were some of the first workers in mass manufacturing
to successfully rally for political causes,
first in the form of a march on parliament in 1871 to protest a proposed tax, and 
later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.

\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
much deeper resentment had been boiling over for years.}
Match--girls were already frustrated with
the arbitrariness of management,
poor working conditions, and
having to work with phosphoric materials (which was proving hazardous).
\ali{I might need to add some info here, and certainly citations, but there'll be overlap with the other citations here}

\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
This was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

\topic{To understand how these women became such influential, if underreported, figures in labor advocacy's history,
we should shine a light onto the reputation that had formed for these workers.}
Match girls were the only group in \citeyear{booth1903life} to have formed a trade union,
according to \citeauthor{booth1903life}'s account at the time~\cite{booth1903life}.
\citeauthor{10.2307/3827491} noted that match girls
``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
These were workers who treasured their independence, but also fiercely protected one another, contributing to the common good.
Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.

\ali{I want to discuss a few things here:
\begin{itemize}
  \item there were protests about something sort of like surge pricing in crop--picking, but i don't know how to make that a paragraph (maybe merge it with match--girls?)
\end{itemize}}

% \topic{When piecework entered the American economy, it was not used for complex work.}
% One reason for low complexity was workers' skills:
% it was infeasible to provide new pieceworkers with the comprehensive education
% that apprenticeships imparted~\cite{hart2013rise}.
% So, initially piecework in the United States arose for farm work, and as
% \citeauthor{hughRaynbirdTaskWork} and others discuss,
% the practice remained relatively obscure until
% it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
% \msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
% The complexity of the work remained low at the turn of
% the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
% \msb{how is this sentence relevant to the case study?:}
% However, writers of the time focused their attention on
% wage~\cite{burton1899commercial} and
% management regimes~\cite{norton1900textile}
% rather than training.
% Mass manufacturing,
% such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
% flourished under piecework systems in densely populated cities.

% \topic{Workers' relationships with employers quickly soured.}
% \msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
% The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
% and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
% As \citeauthor{weyer1894history} described,
% ``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
% the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

% \topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
% \msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
% Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
% Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
% \msb{Why is Taylor relevant to this case study?}
% \ali{I'm just clumsily interleaving the contemporary literature with the events.
% I'd like to say ``This is the geist of work in their time'';
% is this a bad idea?}
% It was in the midst of this time that \citeauthor{taylor1914principles} published
% the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
% It was this framing on work \& worker management that gave workers a concrete adversary
% --- if not Taylor, then \textit{Taylorism}  ---
% against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}

\begin{comment}
What did I pull from the threads that are related to industrial and railroad workers (i.e. 1920 onward?)

- Airy and his human computers were great:
  - quickly verifiable
  - independent tasks (could be checked without the whole product)
  - narrowly trainable

\end{comment}

\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23
\ali{this has a ``gosh isn't that interesting?'' feeling to it.
should i just leave it out until later when there's something to say about it
(namely, that the mode of work was initially designed to stymie professional growth)?}%
.
